---
title: "Project Appendix"
output:
  word_document
---
#Loading Libraries and setting local working Directory
```{r}

library(tseries)
library(fBasics)
library(forecast)
library(lmtest)
library(ggplot2)
library(astsa)
library(fUnitRoots)
library(lubridate)
library(scales) 
library(gridExtra)
setwd("C:\\Users\\guy.dor\\Documents\\CSC 425\\Project")
```

#DAILY DATA

#read data
```{r}
myd=read.table("Speed_Camera_Violations.csv",header=T, sep=',') 
```

#aggregrate data by day
```{r}
smyd <- aggregate(x = myd[c("VIOLATIONS")],FUN = sum,
                  by = list(VIOLATION.DATE = myd$VIOLATION.DATE))
```

#sort data by day
```{r}
dsmyd = smyd[order(as.Date(smyd$VIOLATION.DATE, format="%m/%d/%Y")),]
```
```{r}
head(dsmyd)
basicStats(dsmyd$VIOLATIONS)
```

```{r}
vts= ts(dsmyd[,2], start=c(2014,7,1), freq=7)
```

#histogram
```{r}
hist(dsmyd$VIOLATIONS, xlab="daily violations", prob=TRUE, main="Histogram")
xfit<-seq(min(dsmyd$VIOLATIONS),max(dsmyd$VIOLATIONS),length=5500)
yfit<-dnorm(xfit,mean=mean(dsmyd$VIOLATIONS),sd=sd(dsmyd$VIOLATIONS))
lines(xfit, yfit, col="blue", lwd=2)
lines(density(dsmyd$VIOLATIONS, col= 'blue', lwd =2))
```

#qq plot
```{r}
qqnorm(dsmyd$VIOLATIONS)
qqline(dsmyd$VIOLATIONS, col = 2)
```

#normality test
```{r}
normalTest(dsmyd$VIOLATIONS,method=c("jb")) 
```

```{r}
plot(vts, ylab='Daily Violations', main = 'Daily Speeding Violations')
acf(vts, plot=F) 
acf(vts, plot=T, main = 'ACF on Daily Data') 

pacf(vts, plot=T, main = 'PACF on Daily Data')
```

#Ljung Box Test
```{r}
Box.test(vts, lag= 10, type = 'Ljung')
```

```{r}
adfTest(vts, lags = 180, type = c("c"))
```

# seasonal differencing
```{r}
sdx = diff(vts, 7)
acf(sdx, plot = T)
pacf(sdx, plot = T)
```
```{r}
adfTest(sdx, lags = 180, type = c("c"))
```

#Try automated selection
```{r}
auto.arima(sdx, ic =c("bic"), trace=TRUE)
```

```{r}
m1=Arima(sdx,order=c(0,0,3),seasonal=list(order=c(0,0,1),period=7), method="ML")
m1
```

```{r}
coeftest(m1)
```

#Residual Analysis
```{r}
acf(m1$residuals)
pacf(m1$residuals)
```
```{r}
Box.test(m1$residuals, lag = 10, type = 'Ljung')
```

# found autocorrelation
```{r}
dx = diff(vts)
acf(dx, plot = T)
pacf(dx, plot = T)
```

```{r}
Box.test(dx, lag= 10, type = 'Ljung')
```

```{r}
adfTest(dx, lags = 180, type = c("c"))
```

```{r}
auto.arima(dx, max.p = 10, max.q = 10, ic =c("bic"), trace=TRUE)
```

```{r}
m2=Arima(dx,order=c(1,0,1),seasonal=list(order=c(2,0,0),period=7), method="ML")
m2
```
```{r}
coeftest(m2)
```

#Residual analysis
```{r}
acf(m2$residuals)
pacf(m2$residuals)
```

```{r}
Box.test(m2$residuals, lag = 6, type = 'Ljung', fitdf=4)
```

#Forecast and Backtesting
```{r}
source("backtest.R")
backtest(m2, dx, h=1, orig=length(vts)*0.8)
```

```{r}
m2f=forecast(m2,h=30)
plot(forecast(m2f, h = 30), include = 60)
```
```{r}
m5=Arima(dx,order=c(1,0,7),seasonal=list(order=c(2,0,0),period=7), method="ML")
m5
```
```{r}
coeftest(m5)
```

#Residual Analysis
```{r}
acf(m5$residuals)
pacf(m5$residuals)
```
```{r}
Box.test(m5$residuals, lag = 22, type = 'Ljung', fitdf=8)
```
```{r}
m6=Arima(vts,order=c(1,1,7),seasonal=list(order=c(2,0,0),period=7), method="ML", include.constant = TRUE)
m6
```
```{r}
coeftest(m6)
```

```{r}
acf(m6$residuals)
pacf(m6$residuals)
```

```{r}
Box.test(m6$residuals, lag = 22, type = 'Ljung', fitdf=8)
```

#Forecast and Backtesting
```{r}
source("backtest.R")
backtest(m6, vts, h=1, orig=length(dx)*0.8, inc.drift=TRUE)
```

```{r}
f1=forecast(m6,h=30)
f1
```
```{r}
plot(f1, include =90)
```

#Seasonal differencing
```{r}
sdx7 = diff(dx,7)
acf(sdx7, plot = T)
pacf(sdx7, plot = T)
```

```{r}
adfTest(sdx7, lags = 180, type = c("c"))
```

#Automated selection
```{r}
auto.arima(sdx7, max.p = 10, max.q = 10, ic =c("bic"), trace=TRUE)
```

```{r}
m3=Arima(sdx7,order=c(1,0,0),seasonal=list(order=c(1,0,0),period=7), method="ML")
m3
```

```{r}
coeftest(m3)
```

#Residual Analysis
```{r}
acf(m3$residuals)
Box.test(m3$residuals, lag = 12, type = 'Ljung', fitdf = 2)
```

```{r}
m4=Arima(sdx7,order=c(8,0,0), method="ML")
m4
```
```{r}
coeftest(m4)
```

#Residual analysis
```{r}
acf(m4$residuals)
Box.test(m4$residuals,lag=12,fitdf=8, type='Ljung')
```



#WEEKLY DATA
```{r}
scv=read.table("WeeklyViolations.csv",header=T, sep='\t') 

head(scv)
```
```{r}
x=scv$Violations
head(x)
```
```{r}
xts=ts(x,frequency=52,start=c(2014,7), end = c(2017,8,20))
```

# Create time plot
```{r}
autoplot(xts)+ylab("Speeding Violations")+ggtitle("Weekly Speeding Violations")
```

#Check Normality
```{r}
hist(x,main="Weekly Speeding Violations" , probability = TRUE, xlab = "Speeding Violations")
lines(density(x), col="blue", lwd=2)
lines(density(x, adjust=2), lty="dotted", col="darkgreen", lwd=2) 
```

```{r}
normalTest(x,method=c("jb"))  
```
```{r}
qqnorm(x)
qqline(x, col = 2) 
```

#ACF and PACF of Data
```{r}
acf(x, main="ACF of Weekly Violations")
pacf(x,main="PACF of Weekly Violations")
```

# Dickey Fuller Tests
```{r}
## Tests for AR model with time trend

adfTest(xts, lags=1, type=c("ct"))
adfTest(xts, lags=2, type=c("ct"))
adfTest(xts, lags=4, type=c("ct"))
adfTest(xts, lags=17, type=c("ct"))
adfTest(xts, lags=52, type=c("ct"))

## Tests for AR model with no time trend

adfTest(x, lags=1, type=c("c"))
adfTest(x, lags=2, type=c("c"))
adfTest(x, lags=4, type=c("c"))
adfTest(x, lags=17, type=c("c"))
adfTest(x, lags=52, type=c("c"))

```

#Ljung Box Tests
```{r}
Box.test(x,lag = 1, type="Ljung-Box", fitdf=1)

Box.test(x,lag = 3, type="Ljung-Box", fitdf=2)

Box.test(x,lag = 15, type="Ljung-Box", fitdf=2)
```


#Applying Differencing to Data (s=17)
```{r}
# compute seasonal difference for weeklydata (s=17)
sd17x=diff(x,17)
```

#Create acf plot
```{r}
acf(as.vector(sd17x),lag.max=30, main="ACF of 17 Lag Seasonal Differenced Violations")
pacf(as.vector(sd17x),lag.max=30, main="PACF of 17 Lag Seasonal Differenced Violations")
```

#Ljung Box Tests
```{r}
Box.test(sd17x,lag = 1, type="Ljung-Box", fitdf=1)

Box.test(sd17x,lag = 3, type="Ljung-Box", fitdf=2)

Box.test(sd17x,lag = 15, type="Ljung-Box", fitdf=2)
```

#Check Normality
```{r}
hist(sd17x,main="SDX Lag 17 Speeding Violations" , probability = TRUE, xlab = "SDX Lag 17 Speeding Violations")
lines(density(sd17x), col="blue", lwd=2)
lines(density(sd17x, adjust=2), lty="dotted", col="darkgreen", lwd=2)

qqnorm(sd17x, main="Q-Q Plot on 17 Lag seasonal differenced data")
qqline(sd17x, col = 2)

normalTest(sd17x,method=c("jb"))  
```

#Unit-root tests on Seasonal difference
```{r}
sd17xts=diff(xts,17)
plot(sd17xts)
adfTest(coredata(sd17xts), lags=2, type=c("c"))
adfTest(sd17xts, lags=2, type=c("c"))
```

#Applying Differencing to Data (s=52)
```{r}
sdx=diff(x,52)
```

#Create acf plot
```{r}
acf(as.vector(sdx),lag.max=30, main="ACF of Seasonal Differenced Violations")
pacf(as.vector(sdx),lag.max=30, main="PACF of Seasonal Differenced Violations")
```

#Ljung Box Tests
```{r}
Box.test(sdx,lag = 1, type="Ljung-Box", fitdf=1)

Box.test(sdx,lag = 3, type="Ljung-Box", fitdf=2)

Box.test(sdx,lag = 15, type="Ljung-Box", fitdf=2)
```

#Check Normality
```{r}
hist(sdx,main="SDX Speeding Violations" , probability = TRUE, xlab = "SDX Speeding Violations")
lines(density(sdx), col="blue", lwd=2)
lines(density(sdx, adjust=2), lty="dotted", col="darkgreen", lwd=2)

qqnorm(sdx, main="Q-Q Plot on seasonal differenced data")
qqline(sdx, col = 2)

basicStats(sdx)

normalTest(sdx,method=c("jb"))  
```

#Unit-root tests on Seasonal difference
```{r}
sdxts=diff(xts,52)
plot(sdxts)
adfTest(coredata(sdxts), lags=2, type=c("c"))
adfTest(sdxts, lags=2, type=c("c"))
```

#MODEL EVALUATION OF WEEKLY DATA
# try automated order selection
```{r}
#Original data
auto.arima(xts)
auto.arima(xts,stepwise=FALSE,approx=FALSE)
#Differnced data(s=52)
auto.arima(sdxts)
auto.arima(sdxts,stepwise=FALSE,approx=FALSE)
#Differnced data(s=17)
auto.arima(sd17xts)
auto.arima(sd17xts,stepwise=FALSE,approx=FALSE)
```


```{r}
m0=Arima(xts, order=c(3,1,2),seasonal=list(order=c(1,0,0),period=52), method="ML")
m1=Arima(xts, order=c(2,1,2),seasonal=list(order=c(1,0,0),period=52), method="ML")
```

```{r}
m00=Arima(sdxts, order=c(2,0,0), method="ML")
m11=Arima(sdxts, order=c(1,0,2), method="ML")
```

#m11 can be translated to m12 for the original data
```{r}
m12=Arima(xts, order=c(1,0,2),seasonal=list(order=c(0,1,0),period=52), include.constant=TRUE, method="ML")
```

```{r}
m000=Arima(sd17xts, order=c(3,0,1), method="ML")
m111=Arima(sd17xts, order=c(2,0,1), method="ML")
```

```{r}
m0

m1

m00

m11

m12

m000

m111
```

```{r}
autoplot(m0)
autoplot(m1)
autoplot(m00)
autoplot(m11)
autoplot(m12)
autoplot(m000)
autoplot(m111)
```


#Coefficient tests
```{r}
coeftest(m0)

coeftest(m1)

coeftest(m00)

coeftest(m11)

coeftest(m12)

coeftest(m000)

coeftest(m111)
```


#M1 and (M11) M12 Pass Coef Tests

#Residual Analysis
##M1
```{r}
acf(m1$resid, main="ARIMA(2,1,2)(1,0,0)[52] residual ACF")
pacf(m1$resid, main="ARIMA(2,1,2)(1,0,0)[52] residual PACF")

# ljung box test on residuals
Box.test(m1$residuals, 26, "Ljung-Box", fitdf=5)

Box.test(m1$residuals, 52, "Ljung-Box", fitdf=5)
```

##M12
```{r}
acf(m12$resid, main="ARIMA(1,0,2) with drift residual ACF")
pacf(m12$resid, main="ARIMA(1,0,2) with drift residual PACF")

#  ljung box test on residuals
Box.test(m12$residuals, 26, "Ljung-Box", fitdf=4)

Box.test(m12$residuals, 52, "Ljung-Box", fitdf=4)
```


#FORECASTING
##M1
```{r}
f1=forecast(m1, h=52)

autoplot(f1)+ xlab("Week")+ylab("Weekly Violations")+ggtitle("Forecasts from ARIMA(2,1,2)(1,0,0)[52]")

```

#M12
```{r}
f12=forecast(m12, h=52)

autoplot(f12)+ xlab("Week")+ylab("Weekly Violations")+ggtitle("Forecasts from ARIMA(1,0,2) with drift")

```


#Backtesting
```{r}
source("backtest.R")


backtest(m1, xts,h=1, orig=length(xts)*0.8)


backtest(m12, xts,h=1, orig=length(xts)*0.8)
```




#MONTHLY DATA
```{r}
mcv=read.table("MonthlyViolations.csv",header=T, sep=',') 
head(mcv)
```

```{r}
x=mcv$Violations
head(x)
xts=ts(x,frequency=12,start=c(2014,7))
```

#Create time plot
```{r}
plot(xts, main="Monthly Speeding Violations", xlim=c(2014.5,2017.5), ylim=c(5000,150000),  ylab="Monthly Violations")
autoplot(xts)+ylab("Speeding Violations")+ggtitle("Monthly Speeding Violations")
```

#Check Normality
```{r}
hist(x,main="Monthly Speeding Violations" , probability = TRUE, xlab = "Speeding Violations")
lines(density(x), col="blue", lwd=2)
lines(density(x, adjust=2), lty="dotted", col="darkgreen", lwd=2) 

normalTest(x,method=c("jb"))  

qqnorm(x)
qqline(x, col = 2) 
```

#create acf plots
```{r}
acf(x,lag.max=13, main="ACF Montlhy Violations")
pacf(x, main="PACF Montlhy Violations")
```

# Dickey Fuller Tests
```{r}
# tests for AR model with time trend
adfTest(xts, lags=6, type=c("ct"))
adfTest(xts, lags=12, type=c("ct"))
# tests for AR model with no time trend
adfTest(x, lags=6, type=c("c"))
adfTest(x, lags=12, type=c("c"))
```

#Differencing
```{r}
dx=diff(x)
```

#Create acf plots
```{r}
acf(as.vector(dx),lag.max=13, main="ACF Differenced Violations")
pacf(as.vector(dx),lag.max=13, main="PACF Differenced Violations")
```

#Check for Normality
```{r}
hist(dx,main="DX Speeding Violations" , probability = TRUE, xlab = "DX Speeding Violations")
lines(density(dx), col="blue", lwd=2)
lines(density(dx, adjust=2), lty="dotted", col="darkgreen", lwd=2) 

normalTest(dx,method=c("jb"))  

qqnorm(dx, main="Q-Q Plot on differenced data")
qqline(dx, col = 2)

basicStats(dx)
```

#Unit-root tests on first difference
```{r}
dxts=diff(xts)
plot(dxts)
adfTest(coredata(dxts), lags=2, type=c("c"))
adfTest(dxts, lags=2, type=c("c"))
```

#Seasonal Differencing
```{r}
sdx=diff(x,12)
```

#create acf plotS
```{r}
acf(as.vector(sdx),lag.max=13, main="ACF of Seasonal Differenced Violations")
pacf(as.vector(sdx),lag.max=13, main="PACF of Seasonal Differenced Violations")
```

#Ljung Box Tests
```{r}
Box.test(sdx,lag = 1, type="Ljung-Box", fitdf=1)

Box.test(sdx,lag = 3, type="Ljung-Box", fitdf=2)

Box.test(sdx,lag = 12, type="Ljung-Box", fitdf=2)
```

#Check Normality
```{r}
hist(sdx,main="SDX Speeding Violations" , probability = TRUE, xlab = "SDX Speeding Violations")
lines(density(sdx), col="blue", lwd=2)
lines(density(sdx, adjust=2), lty="dotted", col="darkgreen", lwd=2) 

normalTest(sdx,method=c("jb"))  

qqnorm(sdx, main="Q-Q Plot on seasonal differenced data")
qqline(sdx, col = 2)

basicStats(sdx)
```

#Unit-root tests on seasonal difference
```{r}
sdxts=diff(xts,12)
plot(sdxts)
adfTest(coredata(sdxts), lags=2, type=c("c"))
adfTest(sdxts, lags=2, type=c("c"))
```

# try automated order selection
```{r}
auto.arima(xts)
auto.arima(xts,stepwise = FALSE, approx=FALSE)
```

#Force Seasonal Differencing to compare
```{r}
auto.arima(xts,stepwise = FALSE,D=1, approx=FALSE)
```
```{r}
m3=Arima(xts, order=c(0,1,1),seasonal=list(order=c(1,0,0),period=12), method="ML")
```

#Force Seasonal Differencing to compare
```{r}
m4=Arima(xts, order=c(0,0,0),seasonal=list(order=c(0,1,0),period=12), method="ML")
```


#Coefficient test
```{r}

coeftest(m3)
```
Cannot evaluate m4 coefficients, is a white noise process

#Residual Analysis
```{r}
acf(m3$resid)
pacf(m3$resid)

acf(m4$resid)
pacf(m4$resid)
```

#Forecast

```{r}
f3=forecast(m3, h=12)
f3$mean
autoplot(f3)+ xlab("Month")+ylab("Speeding Violations")+ggtitle("Forecasts from ARIMA(0,1,1)(1,0,0)[12]")
autoplot(f3, xlim=c(2017.65,2018.6))+ xlab("Month")+ylab("Speeding Violations")+ggtitle("Forecasts from ARIMA(0,1,1)(1,0,0)[12]")
```

```{r}
f4=forecast(m4, h=12)
f4$mean
autoplot(f4)+ xlab("Month")+ylab("Speeding Violations")+ggtitle("Forecasts from ARIMA(0,0,0)(0,1,0)[12]")
autoplot(f4, xlim=c(2017.65,2018.6))+ xlab("Month")+ylab("Speeding Violations")+ggtitle("Forecasts from ARIMA(0,0,0)(0,1,0)[12]")
```


#Backtesting
```{r}
backtest(m3, xts,h=1, orig=length(xts)*0.8)
```

```{r}
backtest(m4, xts,h=1, orig=length(xts)*0.8)
```


#ATTEMPTED GARCH

#Violations rate of change
```{r}
VrOc=diff(x)/x[-length(x)]
VrOc
r2=VrOc^2
absr=abs(VrOc)
```

# create acf plot 
```{r}
acf(as.vector(VrOc),lag.max=13, main="ACF of Violations Rate of Change")
pacf(as.vector(VrOc),lag.max=13, main="PACF of Violations Rate of Change")

acf(as.vector(r2),lag.max=13, main="ACF of Violations Rate of Change Squared")
pacf(as.vector(r2),lag.max=13, main="PACF of Violations Rate of Change Squared")

acf(as.vector(absr),lag.max=13, main="ACF of ABS(Violations Rate of Change)")
pacf(as.vector(absr),lag.max=13, main="PACF of ABS(Violations Rate of Change)")
```

#Coefficient test
```{r}
Box.test(VrOc,lag = 1, type="Ljung-Box", fitdf=1)

Box.test(VrOc,lag = 3, type="Ljung-Box", fitdf=2)

Box.test(VrOc,lag = 12, type="Ljung-Box", fitdf=2)
```

#Check Normality
```{r}
hist(VrOc,main="Speeding Violations Rate of Change" , probability = TRUE, xlab = "VrOc")
lines(density(VrOc), col="blue", lwd=2)
lines(density(VrOc, adjust=2), lty="dotted", col="darkgreen", lwd=2) 

normalTest(VrOc,method=c("jb"))  

qqnorm(VrOc, main="Q-Q Plot on Violations Rate of Change")
qqline(VrOc, col = 2)

basicStats(VrOc)
```

#automated selection
```{r}
auto.arima(VrOc)
m5=Arima(VrOc, order=c(0,0,2), method="ML")
m5
```

#Coefficient test
```{r}
coeftest(m5)
```

#Residual Analysis
```{r}
acf(m5$resid)
pacf(m5$resid)
```

#Forecast
```{r}
f5=forecast(m5, h=12)
f5$mean
plot(f5)
```

